---
title: "Portfolio CompMus"
author: "Berend van den Bergh"
date: "30-03-2025"
output: 
  flexdashboard::flex_dashboard:
    storyboard: true
    self_contained: true 
    theme:
      version: 4
      bg: "#101010"
      fg: "#20D20A"
      primary: "#68FF33"
      base_font: 
        google: Prompt
      heading_font:
        google: Sen


---

```{r setup, include=FALSE}
library(tidyverse)
source("compmus.R")
library(flexdashboard)
library(plotly)
library(gridExtra)
library(rvest)
library(tidymodels)
library(ggdendro)
library(heatmaply)
library(patchwork)
```

Introduction {.storyboard}
=========================================


### Welcome to my github page on musicology!

Dear reader, I welcome you to my portfolio! On the next tab you will find some visualisations of the class corpus. 

For this portfolio I have created two songs using AI. The prompts for these songs can be found on the at the bottom of the page. These AI generated songs were inspired by some of the songs featured below. Now, I would like to give you a small challenge: 

Option 1: Have a listen to the songs first and then come up with a prompt, that could make a similar song. After this, compare your prompts to the ones I have used. Did you have something completely different, or something similar?

Option 2: Alternatively, you can start by reading my prompts, then imagine what this type of music would sound like and then listen to the songs below. 

Whichever route you took, be sure to listen to the final result of the AI songs generated by me. Do you think they fit the prompts and/or the songs? They can be found below the prompts at the bottom of the page aswell.


```{=html}

<iframe style="border-radius:12px" src="https://open.spotify.com/embed/track/6iGMRy5RDMiAdTpuMB5gAm?utm_source=generator" width="100%" height="160" frameborder="0" allowfullscreen="" allow="autoplay; clipboard-write; encrypted-media; fullscreen; picture-in-picture" data-external="1"></iframe>

<iframe style="border-radius:12px" src="https://open.spotify.com/embed/track/2I7Ph7hRkjOMgy9WCsE4F3?utm_source=generator" width="100%" height="160" frameborder="0" allowfullscreen="" allow="autoplay; clipboard-write; encrypted-media; fullscreen; picture-in-picture" data-external="1"></iframe>

<iframe style="border-radius:12px" src="https://open.spotify.com/embed?uri=spotify:track:1C5mmdbWD5ksvTU0gHPfao?utm_source=generator" width="100%" height="160" frameBorder="0" allowfullscreen="" allow="autoplay; clipboard-write; encrypted-media; fullscreen; picture-in-picture" data-external="1"></iframe>


```


Here are my AI generated songs. They were created using [StableAudio](https://www.stableaudio.com/)

<details>
  <summary style="color: blue; text-decoration: underline; cursor: pointer;">Show prompts & AI generated songs</summary>
  
The prompt for the first song was: *"A house/hardgroove beat with high energy and danceability, a blend of hefty kicks with catchy riffs and stabs, with a BPM of 143, the key should be in C# Major and the camelot should be 3B."*

cat('<audio controls>
<source src="berend-b-1.mp3" type="audio/mpeg">
Your browser does not support the audio element.
</audio>')


The prompt for the second song was: *"A hardhouse/trance beat with high energy and danceability, with a bpm of 150, the key should be in F# Minor and the camelot should be 11A. It should have a strong basskick."*

My second song

cat('<audio controls>
<source src="berend_b_2.mp3" type="audio/mpeg">
Your browser does not support the audio element.
</audio>')


I came up with these prompts by looking up descriptions of the artists of the songs that I tried to recreate, and copying how their styles were described. I also looked up these songs on [Tunebat](https://tunebat.com/), which is a website that shows lots of musical data on any song via the spotify API.

As a sidenote, [StableAudio](https://www.stableaudio.com/) does have an extensive guide on prompt structure optimalisation, which I did not follow exactly. 

### Visualising the class corpus in a scatterplot




```{r}

classcorp <- read_csv("C:/Users/beren/myrepoMusic/compmus2025.csv")

classcorp <- classcorp |> 
  mutate(highlight = ifelse(filename %in% c("berend-b-1", "berend-b-2"), "yes", "no"))

ggplot(classcorp, aes(x = valence, y = arousal, colour = tempo, size = danceability, 
                          )) + 
  geom_jitter(alpha = 0.8) + 
  ggtitle("Scatterplot displaying the corpus by arousal, valence, tempo and instrumentalness") +
    geom_point(
    data = classcorp |> filter(highlight == "yes"),
    aes(x = valence, y = arousal),
    color = "red",    # Ensures the highlighted songs stand out
    shape = 1,       # Makes them a different shape
    stroke = 0.5,     # Make the outline more visible
  ) +

  # Labels for highlighted songs
  geom_text(
    data = classcorp |> filter(highlight == "yes"),
    aes(label = filename),
    hjust = 0,  
    vjust = -0.6,  
    size = 3.5,       
    color = "red"
  ) +
  scale_colour_viridis_c()    # Use the popular viridis colour palette.

ggplotly()

```

***

This is the entire class corpus, with the valence plotted on the x-axis and arousal on the y-axis. The color is the tempo and the size is the danceability. Interesting to see is that the songs with higher tempo are lower in arousal and danceability. My generated songs show high valence, which can be seen as the musical positivity, as well as high arousal, which can be seen as how much energy the track has. In both these aspects I would agree with these high scores. Lets have a look at how my songs compare to the corpus on some other features.

### Violin plot


```{r}

# Load dataset
classcorp <- read_csv("C:/Users/beren/myrepoMusic/compmus2025.csv")

# Select relevant features
features <- c("approachability", "danceability", "engagingness", "instrumentalness")
filtered_songs <- classcorp %>%
  filter(filename %in% c("berend-b-1", "berend-b-2")) %>%
  select(filename, all_of(features))

# Convert entire corpus to long format
long_data <- classcorp %>%
  pivot_longer(cols = all_of(features), names_to = "variable", values_to = "value")

# Convert selected songs to long format
long_data_songs <- filtered_songs %>%
  pivot_longer(cols = -filename, names_to = "variable", values_to = "value")

# Calculate corpus average
corpus_avg <- classcorp %>%
  summarise(across(all_of(features), mean, na.rm = TRUE)) %>%
  pivot_longer(cols = everything(), names_to = "variable", values_to = "value") %>%
  mutate(filename = "Corpus Average")

# ---- Create Violin Plot ----
ggplot(long_data, aes(x = variable, y = value)) +
  geom_violin(fill = "lightblue", alpha = 0.4) +  # Violin plot for distribution
  geom_jitter(width = 0.2, alpha = 0.2, color = "grey50") +  # Shows all points
  geom_point(data = long_data_songs, aes(color = filename), size = 3) +  # Highlight user songs
  geom_point(data = corpus_avg, shape = 5, color = "black", size = 4) +  # Corpus average (triangle)
  coord_flip() +  # Flip to horizontal for readability
  theme_minimal() +
  labs(title = "Feature Distribution Across Corpus",
       color = "Song") +
  theme(axis.text.x = element_text(size = 12, face = "bold"),
        panel.grid.major = element_line(color = "grey", size = 0.5),
        panel.grid.minor = element_blank())

ggplotly()

```

***

In these violin plots the remaining features can be seen, with my songs highlighted together with the corpus average. It shows that my songs are fairly close to the average in instrumentalness and approachability, but score higher in engaginess and danceability. Instrumentalness predicts the presence of vocals in the song, approachability says something about how "approachable" or "familiar" a song feels. Engaginess measures how compelling or complex the track is. Danceability, like the name implies, measures how well one can dance to the song. 

For most features my two songs are quite close to eachother, except for instrumentalness, and as you might have noticed in the scatterplot, the tempo is quite different. Lets have a closer look at the tempo of my songs.


How different are my two songs? {.storyboard}
=========================================

### Tempograms

```{r, out.width="60%", out.height="60%"}

tempogram1 <- "features/berend-b-1.json" |>
  compmus_tempogram(window_size = 8, hop_size = 1, cyclic = TRUE) |>
  ggplot(aes(x = time, y = bpm, fill = power)) +
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +
  labs(title = "Tempogram song 1", x = "Time (s)", y = "Tempo (BPM)") +
  theme_classic()

tempogram1

tempogram2 <- "features/berend-b-2.json" |>
  compmus_tempogram(window_size = 8, hop_size = 1, cyclic = TRUE) |>
  ggplot(aes(x = time, y = bpm, fill = power)) +
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +
  labs(title = "Tempogram song 2", x = "Time (s)", y = "Tempo (BPM)") +
  theme_classic()

tempogram2


```

***

These are the tempograms of my songs. Looking at the tempogram from song 1, it shows a clear dominant tempo that keeps oscillating around 143 BPM. When listening to the song, it feels fairly repetitive, and the tempogram underlines this clearly. The AI also managed to match the requested BPM of 143 quite exactly.

As for the second song, there is a strong line at around 150 BPM, which also matches the prompt. The two dips in the tempo can be heard in the song. When listening at around 25 seconds and then again at 65 seconds, the "kick" stops for a moment which is causing the tempogram to show a dip in BPM. It is interesting to see that the tempogram picks up on this change, because to me it seems like the overall tempo stays regular in the song. Another interesting detail is that the low-level features measured the BPM to be one tempo-octave lower, that is, 75 BPM.

Lets proceed with our analysis and see if we can see more detail of the song by looking at some novelty functions!


### Novelty Functions

```{r, out.width="100%", out.height="60%"}

energynov1 <- "features/berend-b-1.json" |>
  compmus_energy_novelty() |> 
  ggplot(aes(t, novelty)) +
  geom_line() +
  theme_minimal() +
  labs(title = "Energy Novelty Function Song 1", x = "Time (s)", y = "Energy Novelty")

energynov1

energynov2 <- "features/berend-b-2.json" |>
  compmus_energy_novelty() |> 
  ggplot(aes(t, novelty)) +
  geom_line() +
  theme_minimal() +
  labs(title = "Energy Novelty Function Song 2", x = "Time (s)", y = "Energy Novelty")

energynov2
```

***

This energy novelty function shows that the first song is having little changes in energy except for two peaks. These two peaks are caused by a short silence in the song, and after this silence is broken the energy novelty peaks. It is cool to see that we are basically seeing the same phenomenon as in the tempogram for the second song. 

Looking at the novelty function for the second song, we see a more fluctuating pattern, with different parts highlighted by drops and increases in energy novelty. 

Having seen these novelty functions, we can start to see the differences in the track more. We will now take a look at some of the keys and chords used in the songs.


### Chromagrams

```{r}

p1 <- "C:/Users/beren/myrepoMusic/features/berend-b-1.json" |>                 
  compmus_chroma(norm = "identity") |>                 
  ggplot(aes(x = time, y = pc, fill = value)) + 
  geom_raster() +
  scale_y_continuous(
    breaks = 0:11,
    minor_breaks = NULL,
    labels = c("C", "C#|Db", "D", "D#|Eb", "E", "F", "F#|Gb", "G", "G#|Ab", "A", "A#|Bb", "B")
  ) +
  scale_fill_viridis_c(guide = "none") +               
  labs(title = "Chromagram song 1", x = "Time (s)", y = NULL, fill = NULL) +
  theme_classic()

p2 <- "C:/Users/beren/myrepoMusic/features/berend-b-2.json" |>                 
  compmus_chroma(norm = "identity") |>                 
  ggplot(aes(x = time, y = pc, fill = value)) + 
  geom_raster() +
  scale_y_continuous(
    breaks = 0:11,
    minor_breaks = NULL,
    labels = c("C", "C#|Db", "D", "D#|Eb", "E", "F", "F#|Gb", "G", "G#|Ab", "A", "A#|Bb", "B")
  ) +
  scale_fill_viridis_c(guide = "none") +               
  labs(title = "Chromagram song 2", x = "Time (s)", y = NULL, fill = NULL) +
  theme_classic()

grid.arrange(p1, p2, ncol = 2)  # Zet ncol = 2 als je ze naast elkaar wilt tonen

```

***

To analyze the harmonic structure of my songs lets first have a look at the chromagrams. It is immediately visible that song 1 has mainly G, G# and A#. For the notes from C# to F# there is much less to be seen. For the second song on the other hand, We have a much more even spread throughout the song with leading notes for the first half being F#, C#, which then switches to A# for the second part of the song. 

We can already distinguish a greater difference in tonal centers, so lets have a look at the chordagrams to see if we can find out more about the harmonics of my songs.


### Chordagrams

```{r}

#      C     C#    D     Eb    E     F     F#    G     Ab    A     Bb    B
major_chord <-
  c(   1,    0,    0,    0,    1,    0,    0,    1,    0,    0,    0,    0)
minor_chord <-
  c(   1,    0,    0,    1,    0,    0,    0,    1,    0,    0,    0,    0)
seventh_chord <-
  c(   1,    0,    0,    0,    1,    0,    0,    1,    0,    0,    1,    0)

major_key <-
  c(6.35, 2.23, 3.48, 2.33, 4.38, 4.09, 2.52, 5.19, 2.39, 3.66, 2.29, 2.88)
minor_key <-
  c(6.33, 2.68, 3.52, 5.38, 2.60, 3.53, 2.54, 4.75, 3.98, 2.69, 3.34, 3.17)

chord_templates <-
  tribble(
    ~name, ~template,
    "Gb:7", circshift(seventh_chord, 6),
    "Gb:maj", circshift(major_chord, 6),
    "Bb:min", circshift(minor_chord, 10),
    "Db:maj", circshift(major_chord, 1),
    "F:min", circshift(minor_chord, 5),
    "Ab:7", circshift(seventh_chord, 8),
    "Ab:maj", circshift(major_chord, 8),
    "C:min", circshift(minor_chord, 0),
    "Eb:7", circshift(seventh_chord, 3),
    "Eb:maj", circshift(major_chord, 3),
    "G:min", circshift(minor_chord, 7),
    "Bb:7", circshift(seventh_chord, 10),
    "Bb:maj", circshift(major_chord, 10),
    "D:min", circshift(minor_chord, 2),
    "F:7", circshift(seventh_chord, 5),
    "F:maj", circshift(major_chord, 5),
    "A:min", circshift(minor_chord, 9),
    "C:7", circshift(seventh_chord, 0),
    "C:maj", circshift(major_chord, 0),
    "E:min", circshift(minor_chord, 4),
    "G:7", circshift(seventh_chord, 7),
    "G:maj", circshift(major_chord, 7),
    "B:min", circshift(minor_chord, 11),
    "D:7", circshift(seventh_chord, 2),
    "D:maj", circshift(major_chord, 2),
    "F#:min", circshift(minor_chord, 6),
    "A:7", circshift(seventh_chord, 9),
    "A:maj", circshift(major_chord, 9),
    "C#:min", circshift(minor_chord, 1),
    "E:7", circshift(seventh_chord, 4),
    "E:maj", circshift(major_chord, 4),
    "G#:min", circshift(minor_chord, 8),
    "B:7", circshift(seventh_chord, 11),
    "B:maj", circshift(major_chord, 11),
    "D#:min", circshift(minor_chord, 3)
  )

key_templates <-
  tribble(
    ~name, ~template,
    "Gb:maj", circshift(major_key, 6),
    "Bb:min", circshift(minor_key, 10),
    "Db:maj", circshift(major_key, 1),
    "F:min", circshift(minor_key, 5),
    "Ab:maj", circshift(major_key, 8),
    "C:min", circshift(minor_key, 0),
    "Eb:maj", circshift(major_key, 3),
    "G:min", circshift(minor_key, 7),
    "Bb:maj", circshift(major_key, 10),
    "D:min", circshift(minor_key, 2),
    "F:maj", circshift(major_key, 5),
    "A:min", circshift(minor_key, 9),
    "C:maj", circshift(major_key, 0),
    "E:min", circshift(minor_key, 4),
    "G:maj", circshift(major_key, 7),
    "B:min", circshift(minor_key, 11),
    "D:maj", circshift(major_key, 2),
    "F#:min", circshift(minor_key, 6),
    "A:maj", circshift(major_key, 9),
    "C#:min", circshift(minor_key, 1),
    "E:maj", circshift(major_key, 4),
    "G#:min", circshift(minor_key, 8),
    "B:maj", circshift(major_key, 11),
    "D#:min", circshift(minor_key, 3)
  )

# these functions are not for the chordagrams but for the AI stuff
get_conf_mat <- function(fit) {
  outcome <- .get_tune_outcome_names(fit)
  fit |> 
    collect_predictions() |> 
    conf_mat(truth = outcome, estimate = .pred_class)
}  

get_pr <- function(fit) {
  fit |> 
    conf_mat_resampled() |> 
    group_by(Prediction) |> mutate(precision = Freq / sum(Freq)) |> 
    group_by(Truth) |> mutate(recall = Freq / sum(Freq)) |> 
    ungroup() |> filter(Prediction == Truth) |> 
    select(class = Prediction, precision, recall)
}

```

```{r, out.width="100%", out.height="60%"}

chrdgrm1 <- "features/berend-b-1.json" |> 
  compmus_chroma(norm = "euclidean") |> 
  compmus_match_pitch_templates(
    chord_templates,         # Change to chord_templates if desired
    norm = "euclidean",     # Try different norms (and match it with what you used in `compmus_chroma`)
    distance = "cosine"   # Try different distance metrics
  ) |>
  ggplot(aes(x = time, y = name, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +               # Change the colours?
  labs(title = "Chordagram song 1", x = "Time (s)", y = "Template", fill = NULL) +
  theme_classic()                                      # Change the theme?

#chrdgrm1

chrdgrm2 <- "features/berend-b-2.json" |> 
  compmus_chroma(norm = "euclidean") |> 
  compmus_match_pitch_templates(
    chord_templates,         # Change to chord_templates if desired
    norm = "euclidean",     # Try different norms (and match it with what you used in `compmus_chroma`)
    distance = "cosine"   # Try different distance metrics
  ) |>
  ggplot(aes(x = time, y = name, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +               # Change the colours?
  labs(title = "Chordagram song 2", x = "Time (s)", y = "Template", fill = NULL) +
  theme_classic()                                      # Change the theme?

#chrdgrm2

grid.arrange(chrdgrm1, chrdgrm2, ncol = 2)  # Zet ncol = 2 als je ze naast elkaar wilt tonen


```

***

Here are my two chordagrams. I used a euclidean norm with a cosine distance for both of my songs. Because both of my songs, but the first one especially, are quite repetitive, there is not too much of a prominent chord progression to be seen. In the second song there is a bit of a shift around 50 seconds that is reflected in the chordagram, but for the most part the chords are looping, which is seen more often in the genre of my songs. Perhaps the keygram can tell us more.

### Keygram song 1

```{r}

keygrm1 <- "features/berend-b-1.json" |> 
  compmus_chroma(norm = "euclidean") |> 
  compmus_match_pitch_templates(
    key_templates,         # Change to chord_templates if desired
    norm = "euclidean",      # Try different norms (and match it with what you used in `compmus_chroma`)
    distance = "angular"   # Try different distance metrics
  ) |>
  ggplot(aes(x = time, y = name, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +               # Change the colours?
  labs(title = "Keygram song 1", x = "Time (s)", y = "Template", fill = NULL) +
  theme_classic()                                      # Change the theme?

#keygrm1

keygrm2 <- "features/berend-b-2.json" |> 
  compmus_chroma(norm = "euclidean") |> 
  compmus_match_pitch_templates(
    key_templates,         # Change to chord_templates if desired
    norm = "euclidean",      # Try different norms (and match it with what you used in `compmus_chroma`)
    distance = "angular"   # Try different distance metrics
  ) |>
  ggplot(aes(x = time, y = name, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +               # Change the colours?
  labs(title = "Keygram song 2", x = "Time (s)", y = "Template", fill = NULL) +
  theme_classic()                                      # Change the theme?

#keygrm2

grid.arrange(keygrm1, keygrm2, ncol = 2)  # Zet ncol = 2 als je ze naast elkaar wilt tonen


```


***

The keygram does not show much difference either, as expected since it is also based on harmonics.

Lets move on to another aspect of our songs; the timbre. 



### Cepstrograms

```{r}

cep1 <- "C:/Users/beren/myrepoMusic/features/berend-b-1.json" |>             # Change the track
  compmus_mfccs(norm = "euclidean") |>                  # Change the norm
  ggplot(aes(x = time, y = mfcc, fill = value)) + 
  geom_raster() +
  scale_y_continuous(
    breaks = 0:12,
    minor_breaks = NULL,
  ) +
  scale_fill_viridis_c(guide = "none") +               # Change the colours?
  labs(title = "Cepstrogram song 1", x = "Time (s)", y = "Coefficient Number", fill = NULL) +
  theme_classic()

cep2 <- "C:/Users/beren/myrepoMusic/features/berend-b-2.json" |>             # Change the track
  compmus_mfccs(norm = "euclidean") |>                  # Change the norm
  ggplot(aes(x = time, y = mfcc, fill = value)) + 
  geom_raster() +
  scale_y_continuous(
    breaks = 0:12,
    minor_breaks = NULL,
  ) +
  scale_fill_viridis_c(guide = "none") +               # Change the colours?
  labs(title = "Cepstrogram song 1", x = "Time (s)", y = "Coefficient Number", fill = NULL) +
  theme_classic() 

grid.arrange(cep1, cep2, ncol = 2)  # Zet ncol = 2 als je ze naast elkaar wilt tonen

```


***

The cepstrogram can be used to show the characteristics of the timbre in the song. What we can tell from my cepstrograms is that both songs have a bright sound, since the first coefficient is the most present, and all other coefficients above that are mostly green which suggests a stable and consistent timbre throughout my songs. 

We are getting closer to figuring out how different my songs are. Currently, I would say they are not quite different, rather similar! Perhaps we should look at it from another angle: How similar are my songs?



How similar are my songs? {.storyboard}
=========================================


### Self similarity matrixes

```{r}

ssm1 <- "C:/Users/beren/myrepoMusic/features/berend-b-1.json" |>                   # Change the track
  compmus_chroma(norm = "identity") |>                 # Change the norm
  compmus_self_similarity(
    feature = pc,
    distance = "euclidean"                             # Change the distance
  ) |>   
  ggplot(aes(x = xtime, y = ytime, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +               # Change the colours?
  labs(title = "Self similarity matrix song 1", x = "Time (s)", y = NULL, fill = NULL) +
  theme_classic()                                      # Change the theme?

#ssm1

ssm2 <- "C:/Users/beren/myrepoMusic/features/berend-b-2.json" |>                   # Change the track
  compmus_chroma(norm = "identity") |>                 # Change the norm
  compmus_self_similarity(
    feature = pc,
    distance = "euclidean"                             # Change the distance
  ) |>   
  ggplot(aes(x = xtime, y = ytime, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +               # Change the colours?
  labs(title = "Self similarity matrix song 2", x = "Time (s)", y = NULL, fill = NULL) +
  theme_classic()                                      # Change the theme?

#ssm2

ssm1 + ssm2

#grid.arrange(ssm1, ssm2, ncol = 2)  # Zet ncol = 2 als je ze naast elkaar wilt tonen

```


***

From the self similarity matrixes based on chroma, we a vast repeating pattern for the first song, with strong diagonal lines. At a very small level one can see a bit of a chessboard-like pattern which also makes sense since this first song sounds quite a bit like the same loop repeating over and over again.

For the second song we see more variation, but not as much. Around 50 seconds there is something happening, which corresponds to the beat or "kick" stopping for a while to create tension in the song. 

Lets see if this self similarity is as present in the timbre as well.



### Timbre based SSMs

```{r}

tssm1 <- "C:/Users/beren/myrepoMusic/features/berend-b-1.json" |>                           # Change the track
  compmus_mfccs(norm = "identity") |>                  # Change the norm
  compmus_self_similarity(
    feature = mfcc,
    distance = "euclidean"                             # Change the distance
  ) |>   
  ggplot(aes(x = xtime, y = ytime, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +               # Change the colours?
  labs(title = "Timbre based SSM song 1", x = "Time (s)", y = NULL, fill = NULL) +
  theme_classic()                                      # Change the theme?

#tssm1

tssm2 <- "C:/Users/beren/myrepoMusic/features/berend-b-2.json" |>                       # Change the track
  compmus_mfccs(norm = "identity") |>                  # Change the norm
  compmus_self_similarity(
    feature = mfcc,
    distance = "euclidean"                             # Change the distance
  ) |>   
  ggplot(aes(x = xtime, y = ytime, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +               # Change the colours?
  labs(title = "Timbre based SSM song 2", x = "Time (s)", y = NULL, fill = NULL) +
  theme_classic() 


tssm1 + tssm2
#grid.arrange(tssm1, tssm2, ncol = 2)  # Zet ncol = 2 als je ze naast elkaar wilt tonen


```


***

As expected from looking at the cepstrogram before, the timbre stays very similar, which one can tell by the darkness of the plot. The yellow lines at the edge of the plot indicate the end of the song, where the music is slowly faded out. 

If looked at one by one, my songs show similarity in their progression of chroma and timbre. Lets compare the two songs directly to eachother, both on timbre and chroma.



### Self similarity matrixes: song 1 vs song 2

```{r}

# Load and compute the chroma SSM for both songs
ssm1 <- "C:/Users/beren/myrepoMusic/features/berend-b-1.json" |>  
  compmus_chroma(norm = "identity") |>  
  compmus_self_similarity(feature = pc, distance = "euclidean")

ssm2 <- "C:/Users/beren/myrepoMusic/features/berend-b-2.json" |>  
  compmus_chroma(norm = "identity") |>  
  compmus_self_similarity(feature = pc, distance = "euclidean")

# Compute the absolute difference between the two chroma SSMs
ssm_diff <- ssm1
ssm_diff$d <- abs(ssm1$d - ssm2$d)  # Element-wise difference in similarity

# Plot the difference matrix
ggplot(ssm_diff, aes(x = xtime, y = ytime, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +  
  labs(title = "Difference in Chroma-Based SSM (Song 1 vs. Song 2)", 
       x = "Time (s)", y = "Time (s)") +
  theme_classic()

```

***

Looking at the lot we can finally answer the question how similar or different my songs are. The answer is that they are very similar. Still, they are different songs, but considering they come from the same AI generator, prompted in a very similar way, I am not surprised. Will the timbre based matrix tell a different story?


### Timbre based SSM song 2


```{r}

# Load and compute the SSM for both songs
ssm1 <- "C:/Users/beren/myrepoMusic/features/berend-b-1.json" |>  
  compmus_mfccs(norm = "identity") |>  
  compmus_self_similarity(feature = mfcc, distance = "euclidean")

ssm2 <- "C:/Users/beren/myrepoMusic/features/berend-b-2.json" |>  
  compmus_mfccs(norm = "identity") |>  
  compmus_self_similarity(feature = mfcc, distance = "euclidean")

# Compute the absolute difference between the two SSMs
ssm_diff <- ssm1
ssm_diff$d <- abs(ssm1$d - ssm2$d)  # Element-wise difference in similarity

# Plot the difference matrix
ggplot(ssm_diff, aes(x = xtime, y = ytime, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +  
  labs(title = "Difference in Timbre-Based SSM (Song 1 vs. Song 2)", 
       x = "Time (s)", y = "Time (s)") +
  theme_classic()

```

***

The timbre based self similarity matrix shows that my songs are quite similar too, although there is a bit more difference visible, as for the chroma SSM the pattern seemed more noisy, where as here different lines can still be made out.

Lets check zoom out again and consider the entire class corpus. My songs were both AI-generated, but not the entire class used AI. Lets investigate this further.

As a small sidenote: The way these SSMs are computed is by taking the absolute difference of d, not sure if this is the best way, but I couldnt figure out another way to plot the two songs against eachother.


AI-Insight {.storyboard}
=========================================


### Heatmap Dendogram

```{r}


get_conf_mat <- function(fit) {
  outcome <- .get_tune_outcome_names(fit)
  fit |> 
    collect_predictions() |> 
    conf_mat(truth = outcome, estimate = .pred_class)
}  

get_pr <- function(fit) {
  fit |> 
    conf_mat_resampled() |> 
    group_by(Prediction) |> mutate(precision = Freq / sum(Freq)) |> 
    group_by(Truth) |> mutate(recall = Freq / sum(Freq)) |> 
    ungroup() |> filter(Prediction == Truth) |> 
    select(class = Prediction, precision, recall)
}

compmus2025 <- read_csv("compmus2025.csv")

cluster_juice <-
  recipe(
    filename ~
      arousal +
      danceability +
      instrumentalness +
      tempo +
      valence,
    data = compmus2025
  ) |>
  step_center(all_predictors()) |>
  step_scale(all_predictors()) |> 
  # step_range(all_predictors()) |> 
  prep(compmus2025) |>
  juice() |>
  column_to_rownames("filename")

compmus_dist <- dist(cluster_juice, method = "euclidean")

compmus_dist |> 
  hclust(method = "complete") |> # Try single, average, and complete.
  dendro_data() |>
  ggdendrogram()

heatmaply(
  cluster_juice,
  hclustfun = hclust,
  hclust_method = "average",  # Change for single, average, or complete linkage.
  dist_method = "euclidean"
)

```

***

Before we start looking into if certain songs within the corpus are AI or not, lets first start by visualising some features, to check which ones are useful and which clusters might be present in the corpus.

The dendrogram is quite cluttered, so some zoom is required (this can be done by using your mouse, and double clicking will reset the plot). Still, in the overall picture it is visible that most features have distinct color patterns, which tells us that there is plenty of diversity in the data. Lets see if there are other ways to discover the most important features.


### Random Forest

```{r}

compmus2025_filtered <- 
  compmus2025 |> filter(!is.na(ai)) |> 
  mutate(ai = factor(if_else(ai, "AI", "Non-AI")))

compmus_cv <- compmus2025_filtered |> vfold_cv(5)

classification_recipe <-
  recipe(
    ai ~
      arousal +
      danceability +
      instrumentalness +
      tempo +
      valence,
    data = compmus2025_filtered
  ) |>
  step_center(all_predictors()) |>
  step_scale(all_predictors())      # Converts to z-scores.
  # step_range(all_predictors())    # Sets range to [0, 1].

forest_model <-
  rand_forest() |>
  set_mode("classification") |> 
  set_engine("ranger", importance = "impurity")
indie_forest <- 
  workflow() |> 
  add_recipe(classification_recipe) |> 
  add_model(forest_model) |> 
  fit_resamples(
    compmus_cv, 
    control = control_resamples(save_pred = TRUE)
  )

#indie_forest |> get_pr()

workflow() |> 
  add_recipe(classification_recipe) |> 
  add_model(forest_model) |> 
  fit(compmus2025_filtered) |> 
  pluck("fit", "fit", "fit") |>
  ranger::importance() |> 
  enframe() |> 
  mutate(name = fct_reorder(name, value)) |> 
  ggplot(aes(name, value)) + 
  geom_col() + 
  coord_flip() +
  theme_minimal() +
  labs(x = NULL, y = "Importance")



```

***

In this plot we can see the importance of features that the random forest model is using to determine whether a song is AI or not. Lets plot these variables to show this more clearly. 



### Visualisation of AI or not AI

```{r}

compmus2025_filtered |>
  ggplot(aes(x = valence, y = arousal, colour = ai, size = tempo)) +
  geom_point(alpha = 0.8) +
  scale_color_viridis_d() +
  labs(
    x = "Danceability",
    y = "Instrumentalness",
    size = "Tempo",
    colour = "AI"
  )

ggplotly()
```

***

Plotting these variables shows us clusters of AI and non-AI songs. Knowing what these clustermethods are looking for, lets test their precision and recall.


### KNN, Precision & recall

```{r}

knn_model <-
  nearest_neighbor(neighbors = 1) |>
  set_mode("classification") |> 
  set_engine("kknn")
classification_knn <- 
  workflow() |> 
  add_recipe(classification_recipe) |> 
  add_model(knn_model) |> 
  fit_resamples(compmus_cv, control = control_resamples(save_pred = TRUE))

#classification_knn |> get_conf_mat() |> autoplot(type = "heatmap")
#classification_knn |> get_pr()

indie_forest |> get_pr()

```

```{r}
# Extract precision-recall values
pr_knn <- classification_knn |> get_pr()
pr_forest <- indie_forest |> get_pr()

# Combine results into a table
pr_results <- bind_rows(
  pr_knn |> mutate(Model = "KNN"),
  pr_forest |> mutate(Model = "Random Forest")
)

# Prepare data
pr_results_long <- pr_results |> pivot_longer(cols = c(precision, recall), names_to = "Metric", values_to = "Value")

# Create bar plot
ggplot(pr_results_long, aes(x = class, y = Value, fill = Metric)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~Model) +
  scale_fill_manual(values = c("precision" = "steelblue", "recall" = "darkorange")) +
  labs(title = "Precision and Recall for Classification Models",
       x = "Class",
       y = "Score",
       fill = "Metric") +
  theme_minimal()

ggplotly()
```


***

Here we can see the predictions of the Random Forest and the data on precision and recall for the Random Forest and KNN. 

This data shows what the KNN and Random Forest algorithms predicted and how "accurate" its predictions were. A higher precision means fewer false positives, and a higher recall means fewer false negatives. The plot also shows that the Random Forest performed better overall.


Conclusions {.storyboard}
=========================================

### Conclusion and discussion!

In this project, I analyzed my two tracks using different computational music techniques to better understand their structure and characteristics.

My first song turned out to be quite repetitive, which was clear from its self-similarity matrix and chroma features, showing that it sticks to a loop-based structure, which fits well with house and hardgroove styles. My second song had two drops, and this was reflected in its tempogram, energy novelty function and self-similarity differences, showing a bit more variation and buildup.

Comparing them on harmonics and later by using SSMs I figured out that my songs were, as expected, quite similar to eachother. After this I analyzed the corpus on its distribution of AI and non-AI generated music, and calculated the precision of the Random Forest and KNN algorithms.

Overall, this project helped me get a better grasp of how one can break down music in ways that are not always visible by just listening to the tracks.

As per the discussion:
I am currently following the bachelor AI, and this subject was something entirely new for me. It has proved very difficult for me to understand a lot of the plots and their uses behind it, since I have never played an instrument before I was already struggling with the reading of notes and the many different terms used in music and music theory. Besides this, I had a very different idea of what this course would look like. Most notably for me, was the fact that the music was still very much aimed towards separate instruments and not EDM-style, more house-like genres. 

Nonetheless, it was very cool to see how much there is to music and all the different ways you can compute many different aspects of it. 
