---
title: "Berend's portfolio, homework 8 version"
author: "Berend van den Bergh"
date: "Block 4"
output: 
  flexdashboard::flex_dashboard:
    storyboard: true
    theme: lumen
    self_contained: false
    
---

```{r setup, include=FALSE}
library(tidyverse)
library(flexdashboard)
library(plotly)
```

### Welcome to my github page on musicology!


Dear reader, I welcome you to my github page. On the other tabs you will find some visualisations of the (soon to be class corpus of music) 2024 AI Song contest. 

For this portfolio I have created two songs using AI. The prompts for these songs can be found on the next page. The songs i have created using AI were inspired by some of these songs below. Now, I would like to give you a small challenge: either have a listen to these songs first and then come up with a prompt, to try to make a similar song to the songs below. After this, have a look at the prompts I have used to try and create similar songs. Did these match your idea of the best prompt?

Alternatively, you can take a look at my prompts first, then imagine what this type of music would sound like and then listen to the songs below. 

Whichever route you took, be sure to listen to the final result of the AI songs generated by me. Do you think they fit the prompts and/or the songs?




```{r, results='asis', echo=FALSE}
cat('<iframe style="border-radius:12px" src="https://open.spotify.com/embed/track/6iGMRy5RDMiAdTpuMB5gAm?utm_source=generator" width="100%" height="152" frameborder="0" allowfullscreen="" allow="autoplay; clipboard-write; encrypted-media; fullscreen; picture-in-picture"></iframe>')

cat('<iframe style="border-radius:12px" src="https://open.spotify.com/embed/track/2I7Ph7hRkjOMgy9WCsE4F3?utm_source=generator" width="100%" height="152" frameborder="0" allowfullscreen="" allow="autoplay; clipboard-write; encrypted-media; fullscreen; picture-in-picture"></iframe>')

cat('<iframe style="border-radius:12px" src="https://open.spotify.com/embed/track/1C5mmdbWD5ksvTU0gHPfao?utm_source=generator" width="100%" height="152" frameborder="0" allowfullscreen="" allow="autoplay; clipboard-write; encrypted-media; fullscreen; picture-in-picture"></iframe>')

cat('<audio controls>
        <source src="https://berendvdbergh.github.io/myrepoMusic/berend_b_1.mp3" type="audio/mpeg">
        Your browser does not support the audio element.
     </audio>')

```

### Visualising the class corpus and my own genAI tracks

```{r}
aisc2024 <- read_csv("C:/Users/beren/myrepoMusic/aisc2024.csv")
aisc2024 |>                   # Start with the data
  ggplot(                     # Set up the plot.
    aes(
      x = tempo,
      y = arousal,
      size = instrumentalness,
      colour = danceability
    )
  ) +
  geom_point() +              # Scatter plot.
  geom_rug(linewidth = 0.1) + # Add 'fringes' to show data distribution.
  geom_text(                  # Add text labels from above.
    x = 121,
    y = 4.91,
    label = "Onda Corta - Sud America",
    size = 3,                 # Override size (not loudness here).
    hjust = "left",           # Align left side of label with the point.
    vjust = "center",         # Align vertical centre of label with the point.
    angle = 30                # Rotate the text label
  ) +
  scale_x_continuous(         # Fine-tune the x axis.
    limits = c(50, 200),
    breaks = c(50, 100, 150, 200), # Specify grid lines
    minor_breaks = NULL       # Remove 'minor' grid lines.
  ) +
  scale_y_continuous(         # Fine-tune the y axis in the same way.
    limits = c(1, 9),
    breaks = c(1, 5, 9),
    minor_breaks = NULL
  ) +
  scale_colour_viridis_c() +  # Use the popular viridis colour palette.
  scale_size_continuous(      # Fine-tune the sizes of each point.
    trans = "exp",            # Use an exp transformation to emphasise loud..
    guide = "none"            # Remove the legend for size.
  ) +
  theme_light() +             # Use a simpler theme.
  labs(                       # Make the titles nice.
    x = "Tempo",
    y = "Arousal",
    colour = "Danceability"
  )
ggplotly()
```


***

This is the bad visualisation of the AI Song Contest we used in our first lab session, this time in a dashboard.

### Conclusion and discussion!

Lorum ipsum bzz
